---
title: "p8105_hw2_yg2762"
author: "Yang Gao"
date: "10/1/2021"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

# Problem 1
Read and clean the Mr. Trash Wheel sheet

```{r}
trash_wheel_df= read_excel("./data/trash-wheel.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N406") %>% 
  janitor :: clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate (sports_balls = round(sports_balls, digits= 0))

trash_wheel_df
```


Read and clean precipitation data for 2018 and 2019,omit rows without precipitation data and add a variable for year.
```{r}
precip2019_df = read_excel("./data/trash-wheel.xlsx", sheet="2019 Precipitation", range = "A2:B14") %>% 
  janitor :: clean_names() %>% 
  mutate (year = "2019") %>% 
  drop_na()
  
precip2019_df


precip2018_df = read_excel ("./data/trash-wheel.xlsx", sheet="2018 Precipitation", range = "A2:B14") %>% 
  janitor :: clean_names() %>% 
  mutate (year = "2018") 
  
precip2018_df
```

Combine precipitation datasets and convert month to a character variable

```{r}
precip_df = 
  bind_rows (precip2018_df, precip2019_df) %>% 
  relocate (year)
  
precip_df
``` 
Mr. Trash Wheel data `trash_wheel_df` contains 344 observations and 14 variables; the precipitation dataset`precip_df` contains 18 observations and 3 variables. The total precipitation in 2018 is `sum (precip2018_df$total)`, the median number of sports balls in a dumpster in 2017 is `trash_wheel_filter=filter(trash_wheel_df, year == 2017)
median(trash_wheel_filter$sports_balls )`

```{r}
trash_wheel_df
precip_df

sum (precip2018_df$total)

trash_wheel_filter=filter(trash_wheel_df, year == 2017)
median(trash_wheel_filter$sports_balls )

```

# Problem 2








